---
title: "Sleep Health Analysis"
output: html_document
date: "2024-05-31"
---

Ask 


1. What are some trends in sleep health? 

We can analysis the sleep health to discover patterns affects activity level.



You will produce a report with the following deliverers: 


1. A clear summary of the business task.
 

2.A description of all data sources used. 

4. A summary of your analysis 

5. Supporting visualizations and key findings 

6. Your top high-level content recommendations based on your analysis Use the following 


Ask phase Questions.

What is the problem you are trying to solve? 

Does having bad sleep decrease activities levels compared to the average person in the study.


How can your insights drive business decisions?  

if we could improve the person sleep, we improve health by increasing activity level.

Prepare phase questions.

Where is your data stored? 

One Csv File

How is the data organized?

The data is divided into wide format on daily Activity where each row has observations about the measurable health statistics. 


Does your data ROCCC? 

It's a first hand company acquiring computer data from devices. The devices are computer programmable collections by first hand measurements on the characteristics making them reliable and original. Health data is more timeless due to the nature of how evolution changes very slowly so 8 years ago shouldn’t be a reason to throw it out. 

How are you addressing licensing, privacy, security, and accessibility?

its cited and public domain data. So the data Roccc. The security of the data is a non issue due it containing no personal information that can be used against the clients thanks to its “depersonalized id system”

How did you verify the data’s integrity?  

The data integrity is a little lackluster, its completeness is lacking due to variation in the number of entries for each person suggesting that the data removed null values before publishing the data. The accuracy of the data is only as good as the fitbits natural measurement failures. We are unable to know the life-cycle of how the data was cared for before being submitted on to kaggle for our use. The data integrity is considered low to medium. 

How does it help you answer your question? 

We have a lot activity based data and our goal is discover the changes in the individual over time suggesting to future clients the benefits of using our smart devices. 

The daily activities of these individuals can determine how frequently do our clients use the devices to track 

Their health in order to find out how best to advertise to our clients.

Are there any problems with the data?

The first problem is the stakeholder believes that there are limitations in the data, so we might need to acquire more data to analyze. 

Process 

What tools are you choosing and why? 

I plan on using r due to the large natures of the files in question. 

What steps have you taken to ensure that your data is clean?

Action one: determines if all data in the data set has the same number of observations for every client? 

Answer no each person does not have all the same suggesting the removal during the cleaning process.

Action two: determine the number of observations in each column is the same number of values? 

The data has the same number of numeric values suggesting all the nulls have been removed. 

Action three: determine the rows is the same with numeric values? 

Yes, All data columns have the same number of columns with numeric values in it.

Action four: inspect each column for null values using filters. 


● How can you verify that your data is clean and ready to analyze?

● Have you documented your cleaning process so you can review and share those results? 

Key tasks 

1. Check the data for errors. 

2. Choose your tools. We Choose R

3. Transform the data so you can work with it effectively. 

4. Document the cleaning process. Completed my check 


```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(magrittr)

install.packages("skimr")
library(skimr)
```

```{r}
data <- read.csv('Health_Sleep_Statistics.csv')


```

This is the start of the key task to check the data for errors.
```{r}
skim_without_charts(data)
```
As you can observe the number of n_missing data is zero suggesting all the columns where previously clean for us. 
```{r}
glimpse(data)

```
All the columns have a constant name format therefore do not need to be cleaned. Every word is lowercase and there are no spaces between them removing unnecessary underscores. 

Cleaning the data to confirm that all the columns contain numeric values this step proved to be unnecessary, but it was a fun implementation of code.




Cleaning the data to confirm that all the columns contain numeric values this step proved to be unnessary, but it was a fun implementation of code.

This shows that all the data contains numeric values and not null values suggesting the nulls have been removed from the data prior to our aquiring from kaggle.

checking if all the data contain all the same number of numeric values in each row.



Now lets inspect the data for outliers we need to use a box and wisker plot using the ggplot methodologies.


```{r}
install.packages("ggplot2")  # Install ggplot2 if you haven't already
library(ggplot2)             # Load the ggplot2 package
```

```{r}

plot_boxplot <- function(data, Sleep.Quality) {
  ggplot(data, aes(y = Sleep.Quality)) +
    geom_boxplot() +
    labs(title = paste("Box Plot of", Sleep.Quality),
         y = Sleep.Quality) +
    theme_minimal()
}
plot_boxplot(data, "Sleep.Quality")

```

We observe no outliers
```{r}
col_names <- names(data)
print(col_names)
```

```{r}
col_names <- names(data)

plot_boxplot <- function(data, column_name) {
  ggplot(data, aes_string(y = column_name)) +
    geom_boxplot() +
    labs(title = paste("Box Plot of", column_name),
         y = column_name) +
    theme_minimal()
}

# Get column names
col_names <- names(data)

# Loop through column names and create box plots
for (ele in col_names) {
  # Create a box plot for each column
  print(plot_boxplot(data, ele))
}
```

Coverting Char date column into date format must preform line or break the code
```{r}
data$ActivityDate <- as.Date.character(data$ActivityDate,format = "%m-%d-%Y")
```

The box plots show many outlines, but they are still in the realm of possible answers for a living person. We can throw out the idea the data contains errors of the having to large of a number variety that would throw off our calculations in our analysis. 

That concludes our cleaning process we checked for errors, nulls, and extra large numbers that might throw off our calculations. 


Analyze

Now that your data is stored appropriately and has been prepared for analysis, start putting it to work. Use the following Case


Study Roadmap as a guide:


Guiding questions
● How should you organize your data to perform analysis on it? 
The data will be grouped into sleep quality ranges to be see how the new grouping determines the health statistics suggesting that poor quality sleep affects all areas of life.

● Has your data been properly formatted?

The data is organized perfectly for this type of analysis filtering by person then glimpsing, then visualizing trends to spot the patterns is a valid approach for how the data is formatted. 


● What surprises did you discover in the data?


● What trends or relationships did you find in the data?


● How will these insights help answer your business questions?


1. Aggregate your data so it’s useful and accessible.

2. Organize and format your data. 

3. Perform calculations. 

```{r}

df_filtered <- df %>%
  dplyr::filter(Sleep.Quality <= quantile(Sleep.Quality, 0.25, na.rm = TRUE))

# Create a bar plot of 'Physical.Activity.Level' for the filtered data
ggplot(df_filtered, aes(x = Physical.Activity.Level)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(
    title = "Physical Activity Levels in the Bottom 25% of Sleep Quality",
    x = "Physical Activity Level",
    y = "Count"
  ) +
  theme_minimal()

```

4. Identify trends and relationships. 

A summary of your analysis:


Share

Once you have completed your analysis, create your data visualizations. The visualizations should clearly communicate your

high-level insights and recommendations. Use the following Case Study Roadmap as a guide:
Case Study Roadmap - Share

Guiding questions

● Were you able to answer the business questions? 

The goal of the marketing team was to show how these devices are affecting our clients to reach out to future clients to increase our market share of our product.

● What story does your data tell?

The data shows strong trends across multiple measurements that are measuring different aspects of the life of client showing that the devices information helps people adjust their habits over time to live better life's. 

● How do your findings relate to your original question?

Its a clear advertising total of buy our devices and live a better life. 

● Who is your audience? What is the best way to communicate with them?

The audience is future clients who are unsure the investment in our devices will improve there life. 

● Can data visualization help you share your findings?

The visualizations are what proved the patter that I was looking up until making them, I was unable to spot the strong relationships between the individuals trends and the groups trends are similar in the time series habits of the individuals and group showing increasing the good health habits and decrease of groups. 

● Is your presentation accessible to your audience?

The presentation is for Sršen who want to find out how individuals life's are being altered by the devices to market to future clients. 



